# Air Passenger Time Series
```{r ImportDataAndLibrary}
library(ggplot2)
library(tseries)
library(forecast)
```

```{r}
data<-read.csv("AirPassengers.csv")
head(data)
```


```{r echo=TRUE}
data("AirPassengers")
tsdata<-AirPassengers
#Identifying the class of data
class(tsdata)
```

```{r echo=TRUE}
#Observations of the time series data
tsdata
```

```{r}
boxplot(tsdata,main ="Boxplot for Air Passenegers from 1949 to 1961")
```

```{r}
sum(is.na(data))
```

There do not appear to be any outliers and there are no missing values. Therefore no data cleaning is required.
```{r}
#Check the cycle of data and plot the raw data
cycle(tsdata)
plot(tsdata, ylab="Passengers (1000s)", type="o",main ="Monthly Air Passengers Cycle of data from 1949 to 1961")
```

Every year during the summer months there is an increase flow of passenegers travelling and it decreases after the summer season is over. As a cyclic variation, during the initial months there is a decrease in flow of passengers, then there is a increase in flow of passengers later again it decreases. 

```{r}
tsdata_decom <- decompose(tsdata, type = "multiplicative")
plot(tsdata_decom)
```

The above figure shows the time series decomposition into trend, seasonal and random (noise) . It is clear that the time series is non-stationary (has random walks) because of seasonal effects and a trend (linear trend).

```{r}
boxplot(tsdata~cycle(tsdata),xlab="Date", ylab = "Passenger Numbers (1000's)" ,main ="Monthly Air Passengers Boxplot from 1949 to 1961")
```

In the boxplot there are more passengers travelling in months 6 to 9 with higher means and higher variances than the other months, indicating seasonality with a apparent cycle of 12 months.


